Global optimization method with dual Lipschitz constant estimates for problems with non-convex constraints

### 2020-01-21 2019/NUMTA_Barkalov/Special Issure/article.tex
Move images to main directory
add some references

\begin{keyword}
Global optimization, Multiextremal problems, Non-convex constraints, Lipschitz constant estimates 
\end{keyword}

@@

	The methods for solving the Lipschitzian optimization problems constitute a very important area in the development of the global optimization methods and are the objects of investigations of many researchers. The first algorithms of this class were proposed as early as in the 1970s \cite{Evtushenko1971, Piyavskii1972, Shubert1972, Strongin1970}; since that time, this area has been continuing to develop extensively (see, for example \cite{Evtushenko2009, Evtushenko2013, Strongin2000, Sergeyev2013, Jones2009}). Note that these global optimization algorithms are, as a rule, oriented towards solving either  unconstrained optimization problems 
	\cite{Zilinskas, Sergeyev, Pinter, Jones}
or problems with constraints of a simple structure \cite{Vaz2009, Paulavicius2016}. The solving of the problems with complex non-convex constraints is usually performed with the use of the penalty functions method \cite{Stripinis2019, Pillo2012, Pillo2016}, which has a number of drawbacks.  In particular, this method cannot be applied to the problems with the partially defined functions (i.e. when any constraint is violated, the values of all the other functions of the problem remain undefined).

@@

	Within the framework of our research, an original approach to the minimization of the multiextremal functions with non-convex constraints called the index method of the accounting for the constraints 
	\cite{Strongin} 
was used. This approach is based on separate accounting for each constraint of the problem and does not involve the use of the penalty functions. According to the rules of the index method, every iteration called trial at the corresponding point of the search domain includes a sequential checking of the feasibility of the problem constraints at this point. When the first violation of any constraint is found, the trial is terminated (the values of the remaining problem functions are not computed at this point) and the transition to the next iteration point is initiated. This allows solving the problems with partially defined functions. The experimental comparison of the index method of accounting for the constraints and the penalty function method demonstrated the advantages of the index method in terms of the number of iterations and the number of computations of the problem functions \cite{Barkalov2017_1, Barkalov2017_2}.

@@

\begin{figure}[h]
	\center{\includegraphics[width=0.7\linewidth]{img/epsilon_reserved_solution}}
	\caption{$\varepsilon$-reserved solution}
	\label{fig:eps_reserved_solution}
\end{figure}

@@

	Many well known approaches to solving the multidimensional global optimization problems are based on either explicit or implicit reduction of multidimensional problems to one-dimensional ones (see, for example, the description of methods using the diagonal partitions \cite{Sergeyev2006} or simplicial partitions \cite{Zilinskas2008}). In this paper, we will use the approach based on the dimensionality reduction with the use of the space-filling curve $y(x)$ (the Peano curve). Continuous unambiguous mapping $y(x)$ projecting the interval [0,1] onto an $N$-dimensional hypercube $D$ allows reducing a multidimensional constrained minimization problem in the domain $D$ to a one-dimensional constrained optimization problem in the interval [0,1]
\begin{equation}\label{one_dimensional_problem}
	\varphi(y(x^\ast))=\min \left\{\varphi(y(x)): x \in [0,1], \; g_j(y(x))\leq 0, \; 1 \leq j \leq m\right\},
\end{equation}
The issues of the numerical construction of the Peano space filling curve-type mappings and the corresponding theory were described in detail in \cite{Strongin2000, Sergeyev2013}. Note that the theoretical Peano curve $y(x)$ is a limit object and in practice is replaced by its approximation (called \textit{evolvent}) with predefined precision $2^{-m}$ where the number $m>1$ is a parameter of the construction (the density of the evolvent). In this case, the construction density of the evolvent is determined by the accuracy of the possible practical implementation of the problem solution found. For  illustration, Fig. \ref{fig:evolvent} (a), (b) shows the projections of the interval [0,1] are shown for the mapping $y(x)$ with $N=2$ and $m=4,5$, correspondingly.
\begin{figure}[h]
	\begin{minipage}[h]{0.48\linewidth}
		\center{\includegraphics[width=0.8\linewidth]{img/evolvent_N=2_m=4} \\ a)}
	\end{minipage}
	\hfill
	\begin{minipage}[h]{0.48\linewidth}
		\center{\includegraphics[width=0.8\linewidth]{img/evolvent_N=2_m=5} \\ b)}
	\end{minipage}
	\caption{Evolvent for $N=2$ and $m=4$ (a), $m=5$ (b)}
	\label{fig:evolvent}
\end{figure}

@@

	Step 7. Execute the next trial at the middle point of the interval $(x_{t-1}, x_t)$ if the indices of the points $x_{t-1}$, $x_t$ are not the same, i.e.
$$
	x^{k+1} = \frac{x_t + x_{t-1}}{2}, \; \nu(x_{t-1}) \neq \nu(x_t).
$$
Otherwise, execute a trial at the poin
$$
	x^{k+1} = \frac{x_t+x_{t-1}}{2} + \frac{\mathrm{sign}(z_t-z_{t-1})}{2r_\nu}\left[\frac{\left|z_t-z_{t-1}\right|}{\mu_\nu}\right]^N, \; \nu=\nu(x_{t-1})=\nu(x_t).
$$
When computing the point $x^{k+1}$, use the value of the parameter $r_{\nu}$ fixed at Step 6. The algorithm terminates if the condition $\Delta_{t} \leq \epsilon$ is satisfied where $t$ is from (\ref{R_max_t}) and $\epsilon>0$ is the predefined accuracy.

@@

\begin{figure}[h]
	\begin{minipage}[h]{0.48\linewidth}
		\center{\includegraphics[width=\linewidth]{img/stronginC3_IA} \\ a)}
	\end{minipage}
	\hfill
	\begin{minipage}[h]{0.48\linewidth}
		\center{\includegraphics[width=\linewidth]{img/stronginC3_IA-DL} \\ b)}
	\end{minipage}
	\caption{Solving a problem by IA (a) and IA-DL (b) methods}
	\label{fig:example_strongin}
\end{figure}

>>>>

\begin{keyword}
Global optimization, Multiextremal problems, Non-convex constraints, Lipschitz constant estimates, Dimensionality reduction, Numerical methods
\end{keyword}

@@

	The methods for solving the Lipschitzian optimization problems constitute a very important area in the development of the global optimization methods and are the objects of investigations of many researchers. The first algorithms of this class were proposed as early as in the 1970s \cite{Evtushenko1971, Piyavskii1972, Shubert1972, Strongin1970}; since that time, this area has been continuing to develop extensively (see, for example \cite{Evtushenko2009, Evtushenko2013, Strongin2000, Sergeyev2013, Jones2009}). Note that these global optimization algorithms are, as a rule, oriented towards solving either  unconstrained optimization problems 
	\cite{Paulavicius2014, Sergeyev2015, Pinter1996, Jones1993, Gablonsky2001}
or problems with constraints of a simple structure \cite{Vaz2009, Paulavicius2016}. The solving of the problems with complex non-convex constraints is usually performed with the use of the penalty functions method \cite{Stripinis2019, Pillo2012, Pillo2016}, which has a number of drawbacks.  In particular, this method cannot be applied to the problems with the partially defined functions (i.e. when any constraint is violated, the values of all the other functions of the problem remain undefined).

@@

	Within the framework of our research, an original approach to the minimization of the multiextremal functions with non-convex constraints called the index method of the accounting for the constraints 
	\cite{Strongin2000,Pugliese,Barkalov2002,Strongin2003} 
was used. This approach is based on separate accounting for each constraint of the problem and does not involve the use of the penalty functions. According to the rules of the index method, every iteration called trial at the corresponding point of the search domain includes a sequential checking of the feasibility of the problem constraints at this point. When the first violation of any constraint is found, the trial is terminated (the values of the remaining problem functions are not computed at this point) and the transition to the next iteration point is initiated. This allows solving the problems with partially defined functions. The experimental comparison of the index method of accounting for the constraints and the penalty function method demonstrated the advantages of the index method in terms of the number of iterations and the number of computations of the problem functions \cite{Barkalov2017_1, Barkalov2017_2}.

@@

\begin{figure}[h]
	\center{\includegraphics[width=0.7\linewidth]{epsilon_reserved_solution}}
	\caption{$\varepsilon$-reserved solution}
	\label{fig:eps_reserved_solution}
\end{figure}

@@

	Many well known approaches to solving the multidimensional global optimization problems are based on either explicit or implicit reduction of multidimensional problems to one-dimensional ones (see, for example, the description of methods using the diagonal partitions \cite{Sergeyev2006} or simplicial partitions \cite{Zilinskas2008}). In this paper, we will use the approach based on the dimensionality reduction with the use of the space-filling curve $y(x)$ (the Peano curve). Continuous unambiguous mapping $y(x)$ projecting the interval [0,1] onto an $N$-dimensional hypercube $D$ allows reducing a multidimensional constrained minimization problem in the domain $D$ to a one-dimensional constrained optimization problem in the interval [0,1]
\begin{equation}\label{one_dimensional_problem}
	\varphi(y(x^\ast))=\min \left\{\varphi(y(x)): x \in [0,1], \; g_j(y(x))\leq 0, \; 1 \leq j \leq m\right\},
\end{equation}
The issues of the numerical construction of the Peano space filling curve-type mappings and the corresponding theory were described in detail in \cite{Strongin2000, Sergeyev2013}. Note that the theoretical Peano curve $y(x)$ is a limit object and in practice is replaced by its approximation (called \textit{evolvent}) with predefined precision $2^{-m}$ where the number $m>1$ is a parameter of the construction (the density of the evolvent). In this case, the construction density of the evolvent is determined by the accuracy of the possible practical implementation of the problem solution found. For  illustration, Fig. \ref{fig:evolvent} (a), (b) shows the projections of the interval [0,1] are shown for the mapping $y(x)$ with $N=2$ and $m=4,5$, correspondingly.
\begin{figure}[h]
	\begin{minipage}[h]{0.48\linewidth}
		\center{\includegraphics[width=1.0\linewidth]{evolvent_N=2_m=4} \\ (a)}
	\end{minipage}
	\hfill
	\begin{minipage}[h]{0.48\linewidth}
		\center{\includegraphics[width=1.0\linewidth]{evolvent_N=2_m=5} \\ (b)}
	\end{minipage}
	\caption{Evolvent for $N=2$ and $m=4$ (a), $m=5$ (b)}
	\label{fig:evolvent}
\end{figure}

@@

	Step 7. Execute the next trial at the middle point of the interval $(x_{t-1}, x_t)$ if the indices of the points $x_{t-1}$, $x_t$ are not the same, i.e.
$$
	x^{k+1} = \frac{x_t + x_{t-1}}{2}, \; \nu(x_{t-1}) \neq \nu(x_t).
$$
Otherwise, execute a trial at the point
$$
	x^{k+1} = \frac{x_t+x_{t-1}}{2} + \frac{\mathrm{sign}(z_t-z_{t-1})}{2r_\nu}\left[\frac{\left|z_t-z_{t-1}\right|}{\mu_\nu}\right]^N, \; \nu=\nu(x_{t-1})=\nu(x_t).
$$
When computing the point $x^{k+1}$, use the value of the parameter $r_{\nu}$ fixed at Step 6. The algorithm terminates if the condition $\Delta_{t} \leq \epsilon$ is satisfied where $t$ is from (\ref{R_max_t}) and $\epsilon>0$ is the predefined accuracy.

@@

\begin{figure}[h]
	\begin{minipage}[h]{0.48\linewidth}
		\center{\includegraphics[width=\linewidth]{stronginC3_IA} \\ a)}
	\end{minipage}
	\hfill
	\begin{minipage}[h]{0.48\linewidth}
		\center{\includegraphics[width=\linewidth]{stronginC3_IA-DL} \\ b)}
	\end{minipage}
	\caption{Solving a problem by IA (a) and IA-DL (b) methods}
	\label{fig:example_strongin}
\end{figure}
